---
title: "R HW 4"
author: "David Xiao"
date: "4/6/2021"
output:
  word_document: default
---

*R-Homework 4: Instructions*

    - For this assignment, you will be asked a series of questions that cover the past material we have learned.
    - There are many different ways and functions that you can use to answer the questions!
    - Feel free to use any of these methods/functions to answer the question.



################################
#Question 1- Opening the Dataset
################################

    a) Open either the STAT 200 Data 4.8 XLSX Dataset
    b) Create a original & working copy of the dataset
    
```{r}



#Note: This dataset has been updated and is located on the R Module 4 Canvas Page
```


##################################################
#Question 2- Data Cleaning: Mean/Median Imputation
##################################################

    
    a) Calculate the mean or median of "Scores" variable
```{r}


#Note: Use the "Scores Variable" and not the "Scores.Missing" Variable
```
    
    
    b) Now, do Mean or Median Imputation for the "Scores.Missing" Variable in the Dataset. Use the mean or median value you calculated in 2a to impute. 
```{r}


#Note 1: Make sure to update the Dataset with the Imputed Scores.Missing Variable.

#Note 2: There should be no NAs in the Scores.Missing Variable in the Dataset if you complete 2b correctly
```
    
    c) Reference the top 5 rows for the "Scores.Missing" variable in the dataset.
```{r}


#Note: Case 2 should be a mean or median. Not a NA in the dataset.
```


#############################
#Question 3- Basic Commands 1
#############################

    a) What are the variable names of the Dataset? Use a function to show the variable names and then write them in the quote section below.
```{r}


"Write variable names here"
```

    b) What type of variable is "Scores.Missing", "Sleep", "Note_Type_Name" & "Stat.Anxiety". Write your answer in the quotes below.
```{r}

"Scores.Missing- "
"Sleep- "
"Note_Type_Name- "
"Stat.Anxiety"

```
    
    c) Change the variable "Stat.Anxiety" into a factor variable in the working dataset.
```{r}



```
    
    d) Show how "Stat.Anxiety" is a factor variable in the working dataset now.
```{r}


#Note: Stat.Anxiety should not be listed as a numeric variable
```
   
    
#############################
#Question 4- Basic Commands 2
#############################

    a) Load the "psych" package into R
```{r}

```

    b) Use the describe.by() function with the Stat.Anxiety variable as the factor of interest.
```{r}


#Note: There should be 3 different datasets for each level of Stat.Anxiety
```
    
    
    c) What is the mean for the "Scores.Missing" variable for all participants in the dataset?
```{r}


```


    d) What are the means for the "Scores.Missing" variable for participants with low statistics anxiety, medium statistics anxiety & high statistics anxiety?
```{r}
#Mean for Low Statistics Anxiety



#Mean for Medium Statistics Anxiety



#Mean for High Statistics Anxiety



```


#######################
#Question 5- Subsetting
########################

    a) Find the exam scores for students with low Statistics Anxiety Only 
```{r}

```    
    
    b) Find the exam scores for students with Sleep Scores GREATER than 300
```{r}

```

    c) Find the exam scores for students with Sleep Scores LESS than or EQUAL to 300
```{r}

```

    d) Find the exam scores for students with low statistics anxiety only AND sleep scores GREATER than 300
```{r}


#Note: Output should be- 72 85 95 65 84 84 84 72 83 83 82 82 82 81 80 96 81 89 85 67 73 73 81 79 78 88 77 77 76 76
```
