---
title: "R Homework 5.rmd"
author: "Zizhong David Xiao"
date: "5/01/2023"
output: html_document
---

#################################
#################################
*R Homework 5: Big Data Analysis*
#################################
#################################

In this assignment we will do our first big data analysis with a real dataset!

(Make sure to download the YRBSS Dataset on the Canvas Page)


This assignment will consist of four parts. Reference the R Lesson videos if you have any questions.

- 1. Part 1: Uploading Datasets
- 2. Part 2: Data Cleaning
- 3. Part 3: Data Descriptives
- 4. Part 4: Correlational Analysis
- 5. Part 5: Two-Sample T-Test Analysis


#####################
*Dataset Information*
Youth Risk Behavior Surveillance System (YRBSS)
https://www.cdc.gov/healthyyouth/data/yrbs/index.html (More information here)
#####################

    The Youth Risk Behavior Surveillance System (YRBSS) monitors six categories of health-related behaviors that contribute to the leading causes of death and disability among youth and adults, including—

    - Behaviors that contribute to unintentional injuries and violence
    - Sexual behaviors related to unintended pregnancy and sexually transmitted diseases, including HIV infection
    - Alcohol and other drug use
    - Tobacco use
    - Unhealthy dietary behaviors
    - Inadequate physical activity

    YRBSS also measures the prevalence of obesity and asthma and other health-related behaviors plus sexual identity and sex of sexual contacts.
    
    YRBSS is a system of surveys. It includes 1) a national school-based survey conducted by CDC and state, territorial, tribal, and 2) local surveys conducted by state, territorial, and local education and health agencies and tribal governments. 
    

#UMD Resources (https://www.counseling.umd.edu/)
    
    Counseling Services
    Business Hours: 8:30 a.m to 4:30 p.m. ( Monday through Friday)
    After-Hours Crisis Support Phone Services: 4:30 pm to 8:30 am weekdays, 24 hours/day over the weekend
    Phone Number: (301) 314-7651.
    
    Accessibility and Disability Services
    Business Hours: 8:30 a.m to 4:30 p.m. (Monday through Friday)
    Phone Number: (301) 314-7682.


###########################
###########################
#Part 1: Uploading Datasets
###########################
###########################

Just as in every assignment, we will first load the YRBSS dataset from our downloaded files.
    
#Uploading Dataset
```{r}
library(readxl)


#Note: Get in the habit of always doing a original copy and then working copy.
```



#######################
#######################
#Part 2: Data Cleaning
#######################
#######################

After we uploaded the dataset, now we will to do some data cleaning for it. 


Since this dataset is already very large, we will do "Listwise Deletion" method for cleaning...
```{r}


#Do this on the working dataset
```

How many rows of data did we lose from listwise deletion?
```{r}


#Hint (compare the # rows of the original dataset with the working dataset)
```



############################
############################
#Part 3: Describing Datasets
############################
############################

How many Variables & Rows are there in the dataset? (Provide Code Below & Answer in " ")
```{r}
#Variables
"Insert Answer Here"

#Rows
"Insert Answer Here"
```


Now Check if the Structure of the Variables are Correct in the Dataset
```{r}
#Show Structure of Dataset

```


Convert the Sex Variable from a Character to a Factor Variable in the Working Dataset
```{r}

```


Check if you were able to successful change the Sex Variable into a Factor Variable in the Working Dataset
```{r}

#Note:
    #Female = 1
    #Male = 2
```



#############################
#############################
#Part 4: Correlation Analysis
#############################
#############################

Using the variables in the YRBSS dataset (listed on the Word Document titled YRBSS Variables), select appropriate variables for the correlation analysis:

        - Correlation Analysis: Two Ordinal or Continuous Variables


Feel free to select any variables that you are interested in! We are doing our first real statistical analysis here.

```{r}
#What Two Variables Did you Pick? 
"Insert Answer Here"

#Why did you pick them?
"Insert Answer Here"


```

```{r}
#Make a hypothesis of the relationship between the variables. Try to reference actual Psychological Theory and/or Empirical Research. 
"Insert Answer Here"

#What type of Correlation do you need to run? Pearson or Spearman?
"Insert Answer Here"

```

Describe the Variables you Selected
```{r}
#Are these two variables continuous or ordinal?
"Insert Answer Here"


#What is the Mean & Median for Variable 1?
"Insert Answer Here"


#What is the Mean & Median for Variable 2?
"Insert Answer Here"
```

######################
*Statistical Analysis*
######################

Now let's use the cor.test() function to test our correlational analysis.
```{r}


#Note: Make sure to designate pearson or spearman... If you do not, pearson will be the default.
```

Was your hypothesis supported? How do you know?
```{r}
"Insert Answer Here"

#Note: Reference the p-value and directionality of the relationship
```

Provide an explanation for why your hypothesis was supported or not supported. Again try to reference psychological theory or any possible limitations of the survey design and/or sample of the dataset.
```{r}
"Insert Answer Here"


```



##################################
##################################
Part 5.	Two-Sample T-Test Analysis
##################################
##################################

Using the variables in the YRBSS dataset (listed on the Word Document titled YRBSS Variables), select appropriate variables for the correlation analysis:

        - Two-Sample T-Test Analysis: One Binary Variable & One Continuous or Ordinal Variable

Again, feel free to select any variables that you are interested in!


```{r}
#What Two Variables Did you Pick?
"Insert Answer Here"

#Why did you pick them?
"Insert Answer Here"

```

```{r}
#Make a hypothesis of the relationship between the variables. Try to reference actual Psychological Theory and/or Empirical Research. 
"Insert Answer Here"

```

Describe the Variables you Selected
```{r}
#First, Make sure that your binary factor is a factor variable! Use levels() to check if you were able to successfully convert the variable to a factor variable.


```

```{r}
#What is the sample size (i.e., # of rows) for each of level of the Binary Factor Variable?
"Insert Answer Here"


#Hint: try Googling the table() function to do this... This is a shortcut! Otherwise you can try other methods we have learned to do so. There are many ways to do this!
```

```{r}
#What is the Mean & Median for the  Continuous or Ordinal Variable?
"Insert Answer Here"


#Is this variable Continuous or Ordinal?
"Insert Answer Here"

```

######################
*Statistical Analysis*
######################

Now let's use the t.test() function to test our two-sample t-test. Let's make our t-test two-sided and assume that the variance of both groups are equal (i.e. pooled variance estimate or student's t-test)!

Remember X will be our binary factor variable and y will be our ordinal/continuous variable.
```{r}


#Hint: t.test(y variable ~ x variable, alternative = "two.sided or one.sided", var.equal = TRUE or FALSE)
#Hint: t.test(y variable ~ x variable, alternative = "two.sided", var.equal = TRUE)
```

Was your hypothesis supported? How do you know?
```{r}
"Insert Answer Here"

#Note: Reference the p-value and directionality of the relationship (reference the means)
#Note: Remember what group 1 & group 2 refer to!
```

Provide an explanation for why your hypothesis was supported or not supported. Again try to reference psychological theory or any possible limitations of the survey design and/or sample of the dataset.
```{r}
"Insert Answer Here"


```
