---
title: "R Interactive Video 5"
author: "Zizhong (David) Xiao"
date: "4/28/2023"
output: word_document
---

###############################
###############################
*Video 1_R Lesson 6: Big Data Analysis*
###############################
###############################

In this lesson, we will integrate all of the material from our prior lessons and then apply them to analyze a real-world dataset!

- 1. Part 1: Uploading Datasets
- 2. Part 2: Data Cleaning
- 3. Part 3: Data Descriptives
- 4. Part 4: Data Analysis (Correlation & T-Tests)


###################
*Dataset Information*
Youth Risk Behavior Surveillance System (YRBSS)
https://www.cdc.gov/healthyyouth/data/yrbs/index.html (More information here)
###################

    The Youth Risk Behavior Surveillance System (YRBSS) monitors six categories of health-related behaviors that contribute to the leading causes of death and disability among youth and adults, including—

    - Behaviors that contribute to unintentional injuries and violence
    - Sexual behaviors related to unintended pregnancy and sexually transmitted diseases, including HIV infection
    - Alcohol and other drug use
    - Tobacco use
    - Unhealthy dietary behaviors
    - Inadequate physical activity

    YRBSS also measures the prevalence of obesity and asthma and other health-related behaviors plus sexual identity and sex of sexual contacts.

    YRBSS is a system of surveys. It includes 1) a national school-based survey conducted by CDC and state, territorial, tribal, and 2) local surveys conducted by state, territorial, and local education and health agencies and tribal governments. 
    

#UMD Resources (https://www.counseling.umd.edu/)
    
    Counseling Services
    Business Hours: 8:30 a.m to 4:30 p.m. ( Monday through Friday)
    After-Hours Crisis Support Phone Services: 4:30 pm to 8:30 am weekdays, 24 hours/day over the weekend
    Phone Number: (301) 314-7651.
    
    Accessibility and Disability Services
    Business Hours: 8:30 a.m to 4:30 p.m. (Monday through Friday)
    Phone Number: (301) 314-7682.


###########################
###########################
#Part 1: Uploading Datasets
###########################
###########################

We will use the YRBSS Dataset that is located in our files section. This dataset is just a subset of the entire YRBSS Dataset.

Sample Size of 14,765! (This is Relatively Huge for Social Science Research)

Reference the attached Word File (YRBSS Variables) on the class module page to see how each of these variables were scored.

Variables:
      
      ID (Participant Number)
      Age
      Sex Assigned at Birth (Male/Female)
      
      How Often Participant Carries a Gun
      How Often Participant Carries a Gun at School
      
      How Often Participant Feels Unsafe on way to School
      How Often Participant is Threatened at School
      
      If participant has been forced to have sexual intercourse
      How many times Participant has been victim of any sexual violence
      
      If Participant has been Bullied at School
      If Participant has been Bullied Online
      
      If Participant Considered Suicide
      If Participant Made Suicide Plan
      How many times Participant Attempted Suicide
      
      How Often Participant Drinks Alcohol (per Week)
      How Often Participant Binge Drinks (per Month)
      Most Drinks of Alcohol Participant has had in one day (per Month)
      
      How many times Participant used Cocaine
      How many times Participant used Inhalant
      How many times Participant used Heroin
      How many times Participant used Meth
      How many times Participant used Ecstasy
      How many times Participant used Steroid
      How many times Participant used Ever Abused Prescription Drugs
      
    
#Uploading Dataset
```{r}
library(readxl)
dat.org <- read_xlsx("YRBSS_Psyc200.xlsx")
#working copy
dat <- dat.org
```


#######################
#######################
#Part 2:	Data Cleaning
#######################
#######################

After we uploaded the data,  we will to do some data cleaning for it. 

Since this dataset is already very large, we will do "Listwise Deletion" method for cleaning. 
```{r}
dat.clean<-na.omit(dat)
```

However, notice that we lose nearly 50% of our data here! So there are negatives too that come with list-wise deletion.
```{r}
nrow(dat)
nrow(dat.clean)
```

In most cases, mean imputation is the default (common) method for data cleaning. BUT, it is not appropriate for factor variables nor binary variables.

    - Ex: Mean Imputation Inappropriate for...
      - Gender at Birth (Binary-Factor)
      - Forced Sex (binary)
      - Bullied at School (binary)
      - Bullied Online (binary)
      - Suicide Consider (binary)
      - Suicide Plan (binary)
      

############################
############################
#Part 3: Describing Datasets
############################
############################

#Basic Dataset Functions
```{r}
#Dimensions (7874 Rows Long, 25 Variables)
dim(dat.clean)

#Variables
colnames(dat.clean)
```

#Check Datasets if Structures are Correct
```{r}
#Structure
str(dat.clean)
```

Here we see that Sex is a Character Variable. We will want to change these variables.

    - Sex --> Factor Variable

```{r}
#Convert Sex to Factor (please pay attention to this process, many of you had questions)
dat.clean$Sex <- as.factor(dat.clean$Sex)

#Checking if Variable now a Factor
str(dat$Sex) #Original
str(dat.clean$Sex) #New
    #Female = 1
    #Male = 2


```

Describe Dataset
```{r}
library(psych)
describe(dat.clean)
```


#######################################
#######################################
Part 4.	Conducting Statistical Analyses
#######################################
#######################################

##############
*Correlations*
##############

#Correlation Function
```{r}
?cor.test()
```

*Pearson Correlation*
- 2 Continuous Variables
    - X: How many times Participant has been victim of any sexual violence
    - y: How many times participant has used cocaine
```{r}
#Pearson
cor.test(dat.clean$Sex_Violence, dat.clean$Ever_Cocaine, method = "pearson")


###Results###
#Significant Correlation
#Positive Weak Correlation
#Increases in Sex Violence positively associated with more uses of Cocaine

###Formula###
#cor.test(x, y, method = "name")
```

*Spearman Correlation*
- If at least 1 or both variables are ordinal
- More appropriate technically speaking than Pearson Correlation for last analysis as the Threat at School & Cocaine use variables are both ordinal in nature (they do notincrease at a constant rate)


```{r}
#Spearman
cor.test(dat.clean$Sex_Violence, dat.clean$Ever_Cocaine, method = "spearman")

###Results###
#Significant Correlation
#Positive Weak Correlation
#Increases in Sex Violence positively associated with greater cocaine use, even after using spearman instead of Pearsons
#Slightly less (but in theory more accurate) correlation than the previous question! .22 --> .12

###Formula###
#cor.test(x, y, method = "name")
```


###################
*Video 2a_One-Sample T-Test*
###################

According to drugrehab.com, full-time college students drink on average of 6.4 days per month (https://www.drugrehab.com/2016/06/07/college-student-drinking-statistics/)

In this test we will see if this value (6.4) is significantly different than the YRBSS participants (middle schoolers & high schoolers) - the concept of one-sampled T-test.

#Alcohol Days

    During the past 30 days, on how many days did you have at least one drink of alcohol? 
      1.	0 days 
      2.	1 or 2 days 
      3.	3 to 5 days 
      4.	***6 to 9 days (where 6.4 is located)***
      5.	10 to 19 days 
      6.	F. 20 to 29 days 
      7.	G. All 30 days 

```{r}
?t.test()
#t.test(x, mu = null value, alternative = "one or two-sided") - directional or non-directional?
```

Are YRBSS students drinking habits per month different than full-time college students?
```{r}
t.test(dat.clean$Alcohol_Days, mu = 4, alternative = "two.sided")
#Note: mu = 4, because 6.4 drinks is between 6 to 9

###Results###
#Significant One-Sample T-Test
#Mean of X = 1.47 (vs. 4)
#YRBSS Participants drink significantly less than Full-time College Students
```


###################
*Video 2b_Two-Sample T-Test*
###################

- Also referred to as an Independent Samples T-Test
- Occur when we are compare the mean of some DV between two different groups

In this analysis we will compare whether the mean of suicide attempts is significantly different for YRBSS participants who have been bullied at school.

#Bullied at School (IV) - the independent (grouping) variable

    During the past 12 months, have you ever been bullied on school property? 
      Yes (2)
      No  (1)

#Suicide Attempt (DV) - the dependent variable

    During the past 12 months, how many times did you actually attempt suicide? 
    1.	0 times 
    2.	1 time 
    3.	2 or 3 times
    4.	4 or 5 times
    5.	6 or more times 


Same function but we put different inputs in here! 
```{r}
?t.test()

#t.test(y variable ~ x variable, alternative = "two.sided or one.sided", var.equal = TRUE or FALSE)

#var.equal = TRUE --> Pooled Variance Estimate (equal variance)
#var.equal = FALSE --> Unpooled Variance Estimate (unequal variance)
#In this case, the variance is FALSE.
```

Do suicide attempts differ for participants who have been bullied or not at school?
```{r}
t.test(dat.clean$Suicide_Attempt ~ dat.clean$Bullied_School, alternative = "two.sided", var.equal = FALSE)
###Results###
#Significant Two-Sample T-Test
#Suicide Attempts significantly differ for participants who have been bullied at school
#Suicide Attempts greater in group 2 (bullied at school) than group 1 (not bullied at school). 

```

