---
title: "R Interactive Video 4"
author: "Zizhong Xiao"
date: "3/29/2021"
output: word_document
---

############################
############################
*Video 1. Review of Early Lessons*
############################
############################

- We will start by first opening the STAT 200 Data Update - 4.8 XLSX file in the dataset. 
- Remember to create a original file and a working file


#################
#Opening Datasets
#################

#XLSX File
- We will start by opening the XLSX Dataset
- We will make original copies and working copies
```{r}
library(readxl)
data.original1 <- read_xlsx("STAT200 Data Update - 4.8.xlsx")
data.working <- data.original1
```


##############################
Structure & Summary of Dataset
##############################
```{r}
#str() allows us to see the structure of the variables in the dataset
str(data.working)

#summary provides us with a summary of the variables in the dataset
summary(data.working)
```


##################################
#Describe / Describe.By() Datasets
##################################

#Be sure to have library(psych) to run the describe() functions
```{r}
library(psych)
```


#describe()
```{r}
#describe() also provides us a clean summary/information on the variables in the dataset
  #describe(dataset)
describe(data.working)

#Note: More information provided on describe function here: https://www.rdocumentation.org/packages/psych/versions/2.0.12/topics/describe
```

#describe.by()
```{r}
#describe.by() provides us a clean summary/information on the variables in the dataset FOR A SPECIFIC GROUP

#I am interested in the scores, by ANXIETY group

################################
#Preparing the Data into Factors
################################

#Original Dataset Structure
str(data.original1)

#The orignal data for anxiety is "numbers", but I want it to be a "factor"

#Converting Stat Anxiety from Numeric to Factors
data.working$Stat.Anxiety_Name<-as.factor(data.working$Stat.Anxiety_Name)
  #Low Anxiety
  #Medium Anxiety
  #High Anxiety

#Observing Change from Numeric to Factor
str(data.working)


#################
#Running the Code
#################
#describe(dataset, dataset$group)
describe.by(data.working, data.working$Stat.Anxiety_Name)
#Note2: More information provided here: https://www.rdocumentation.org/packages/psych/versions/2.0.12/topics/describeBy
```


#####################################
#####################################
*Video 2. Review of R Lesson 4: Subsetting*
#####################################
#####################################

#Selecting specific Variables in the Dataset
```{r}
#This is how to Extract a specific variable via $
data.working$Sleep
data.working$Motivation
  #dataset$variable

#This is how to Extract a specific variables (columns) via []
data.working[,2]
data.working[,3]
data.working[,2:3]
  #dataset[row,column]

#This is how to Extract a specific participants (row) via []
data.working[1,]
data.working[3,]
data.working[1:3,]

#This is how to Extract a specific variable AND row via []
data.working[1,1] #row 1 and column 1
data.working[1:3, 1:3] #row 1 to 3 & column 1 to 3
```


###############
*Subsetting*
###############

- Here we look at exam scores by each factor of Statistics Anxiety by directly referencing variables we are interested in using [ ] and logical operators

#Factors
- Here we subset for Factors
```{r}
#Exam Scores for Students with Low Stat Anxiety (Stat Anxiety == 1)
data.working$Scores[data.working$Stat.Anxiety == 1]
#Exam Scores for Students with High Stat Anxiety (Stat Anxiety == 3)
data.working$Scores[data.working$Stat.Anxiety == 3]
```

#Numeric
- Here we subset for numerical information
```{r}
#Exam Scores for Students with Motivation Equal to 5
data.working$Scores[data.working$Motivation ==5]
#Exam Scores for Students with Motivation Greater than 5
data.working$Scores[data.working$Motivation > 5]
#Exam Scores for Students with Motivation Less than 5
data.working$Scores[data.working$Motivation < 5]
```

#######################
#AND Logical Operator &
#######################
```{r}
#Exam Scores for Statistics Anxiety Low (1) & Motivation Greater than 5
data.working$Scores[data.working$Stat.Anxiety == 1 & data.working$Motivation > 5]

data.working$Scores[data.working$Stat.Anxiety == 1 & data.working$Sleep > 300]

#Exam Scores for Statistics Anxiety High (3) & Motivation Less than or Equal to 5
data.working$Scores[data.working$Stat.Anxiety == 3 & data.working$Motivation <=5]
```



######################
######################
*Video 3. Cleaning Datasets*
######################
######################

Much of work in RStudio and data analysis involves processing datasets. This is not the glamorous part of statistical analysis but it makes up more than 75% of the job itself. In this section, we will discuss three popular methods for data cleaning.


First, let's check the Total Sample Size of Dataset
```{r}
#Sample Size Original/Working Datasets
nrow(data.original1)
nrow(data.working)
#as we can see, N = 90 as we have 90 participants in total
```

Now, lets reference the variables that have missing data to them. Here when we reference the missing_data variable, we can see that we are missing some values below
```{r}
data.working$Stat.Anxiety
data.working$Stat.Anxiety.Missing
#Note: The Missing Data is referenced as NA
```

*The missing data is very common and may be due to:*
    
    - Participant Fatigue
    - Privacy Concerns
    - Connectivity Issues
   

For data analysis, the easiest way to address missing values consist of three methods:

    1) Listwise Deletion
    2) Casewise Deletion
    3) Mean/Median Imputation

Now, we will discuss these three methods.


###################
*Listwise Deletion*
###################

    - This method refers to deleting all rows in the dataset that contain at least one missing data value in all of the variables
    - na.omit() can be used to limit data to only rows with data for all variables
    - One Problem with this method is that it can eliminate a lot of data

```{r}
data.listwise.deletion <- na.omit(data.working)
data.listwise.deletion #Notice the Missing_Data Column
View(data.listwise.deletion)

#Compare Listwise & Working
head(data.original1$Stat.Anxiety.Missing)
head(data.listwise.deletion$Stat.Anxiety.Missing)

nrow(data.original1) #N = 90
nrow(data.listwise.deletion) #N = 67
```


###################
*Casewise Deletion*

    - This method refers to just ignoring the data for a certain type of analysis when it is missing
    - R requires us to do this for many analyses by ignoring the data that is missing a value; but also does it automatically for some analyses by ignoring the data that is missing a value.
    

#Example where R Does not do it Automatically

    - To specify casewise deletion manually, you can input a na.rm = TRUE within the function itself.

```{r}
#R Does not Work
mean(data.working$Stat.Anxiety.Missing)
median(data.working$Stat.Anxiety.Missing)


#R Works when we do it Manually
mean(data.working$Stat.Anxiety.Missing, na.rm = TRUE)
median(data.working$Stat.Anxiety.Missing, na.rm = TRUE)
```


#Example where R Does it Automatically

    - However, sometimes R functions do casewise deletion automatically!
    - The cor.test() function does this!
    - Unfortunately, we have to check each function to see how it treats missing data. 


#Example where R does Casewise Deletion Automatically
```{r}
#Correlation between Statistics Anxiety (Missing) & Exam Scores
cor.test(data.working$Stat.Anxiety.Missing, data.working$Scores)
    #cor = -.14
    #df = 74

```


########################
*Median/Mean Imputation*

    - This method refers to filling in the missing data values with the mean or median of said variable
    - For instance if the mean is 5 for motivation variable, we would fill in all of the NAs for motivation variable with 5's
    - The benefits of mean/median imputation is that it allows us to have more data to use for our analyses
    - The major downside is that it reduces our variability in the dataset
    
    
Preparing Data for Mean/Median Imputation
```{r}
#Create a Copy for Mean Imputation
mean_impute <- data.working
median_impute <- data.working

#Calculate Mean/Median for Stat Anxiety, using casewise deletion
mean(mean_impute$Stat.Anxiety.Missing, na.rm=TRUE)
median(median_impute$Stat.Anxiety.Missing, na.rm = TRUE)
    #Mean = 1.986842
    #Median = 2

```

#Doing Mean Imputation

```{r}
################
#Mean Imputation
################

"NameofDataset$VariabletoReference[is.na(#nameofDataset$VariabletoReference)] <- ValuetoReplaceMissingData"

#Mean Imputation
mean_impute$Stat.Anxiety.Missing[is.na(mean_impute$Stat.Anxiety.Missing)]<-1.9868642

#Observing Change
mean_impute$Stat.Anxiety.Missing

#Compare to Old
data.working$Stat.Anxiety.Missing

#Note: Significant figures also now change since we added a bunch of figures when we added the mean (e.g., 1.9868642)
```


Doing Median Imputation

    - Same format as Mean Imputation, except now we just use 2 instead of 1.9868642
    
```{r}
##################
#Median Imputation
##################

"NameofDataset$VariabletoReference[is.na(#nameofDataset$VariabletoReference)] <- ValuetoReplaceMissingData"

#Mean Imputation
median_impute$Stat.Anxiety.Missing[is.na(median_impute$Stat.Anxiety.Missing)] <- 2

#Observing Change
median_impute$Stat.Anxiety.Missing

#Compare to Old
data.working$Stat.Anxiety.Missing

#Note: Significant figures do not chance since we did not add extra figures after adding the median (e.g, 2)
```
