---
title: "Lesson 3"
author: "Jordan Epistola"
date: "2/16/2021"
output: html_document
---

#################
*Important Note!*
#################

- After you run the install.packages("") function, make sure to put a # before it or delete the line to ensure that the code does not run again

- If you try to knit with the install.packages("") function again, then it will run into problems as RStudio will have trouble installing the package again, since it is already installed.


###########################################
*Accessing Datasets from R-Studio Packages* 
###########################################

- We will start this lesson by accessing some datasets from the "datasets" package we used in Lesson 1.
- Again, if you already installed this package into R you do not need run the install.packages() line again

```{r}
#install.packages("datasets")
library(datasets)

```



#ChickWeight & VADeaths Dataset
We will access these two datasets below in this package

    - ChickWeight = Weight versus age of baby chickens on different diets
    - VADeaths = Dataset of Death Rates in Virginia (1940)


Good Practice/General Recommendation for Loading Datasets into R
```{r}
#Original Dataset
ChickWeight
VADeaths

#Dataset we work with
dat.working1 <- ChickWeight
dat.working2 <- VADeaths

#Viewing Dataset
dat.working1
dat.working2
```


###################################
*Opening Datasets from Local Files*
###################################

#Installing Packages to Load Datasets

    - To open up excel files and csv files we first need to install and then open a package that allows us to do
    - Below are popular packages for opening excel & Csv files
    
```{r}
#Excel Datasets
#install.packages("readxl")
library(readxl)

#CSV Datasets
#install.packages("readr")
library(readr)


#Note: Remember to delete the install.packages() line or put a # in front of it 
```

Opening Excel Dataset

    - If the datafile is located in the same folder you can just type the name of the dataset in " "
    - If the datafile is located in a different folder, you can find it by typing ../ and selecting enter

    - For this assignment we will work on a STAT200 Dataset which examines test scores and variety of other predictors such as:
          - How much sleep students got
          - If students used computers or paper to write notes

Datasets can be in different formats. The most common forms are 'xlsx' and 'csv.'
```{r}
# This is an example -- you will need to change the green text to the place where the data file is located ON YOUR OWN COMPUTER

Excel.dat.orginal <- read_xlsx("STAT200 Data.xlsx")
Excel.dat.orginal <- read_xlsx("../Stat 200 - Spring 2021/STAT200 Data.xlsx")

Excel.dat.working <- Excel.dat.orginal

Excel.dat.working
```

Opening CSV Dataset
```{r}
CSV.dat.orginal <- read.csv("STAT200 Data.csv")
CSV.dat.orginal <- read.csv("../Stat 200 - Spring 2021/STAT200 Data.csv")

CSV.dat.working <- CSV.dat.orginal

CSV.dat.working
```

Make sure the excel & csv files match the function that is chosen! 
If you have an xlsx make sure to use the read_xlsx()
```{r}
#error1 <- read_xlsx("STAT200 Data.csv")

```

#########################
*Basic Dataset Functions* VIDEO 4
#########################

In this section we will show several basic dataset commands using the PSYC200 Dataset we loaded into R. 
Let use the Excel File (Excel.dat)

Viewing Dataset
```{r}
#View Entire Data
View(Excel.dat.working)

#Top 5 & Bottom 5
head(Excel.dat.working)
tail(Excel.dat.working)
```

Structure & Summary of Dataset
```{r}
#str() allows us to see the structure of the variables in the dataset
str(Excel.dat.working)

#summary provides us with a summary of the variables in the dataset
summary(Excel.dat.working)

```

Descriptives of Dataset
```{r}
#Number of Rows & Columns
nrow(Excel.dat.working)
ncol(Excel.dat.working)

#Column names
colnames(Excel.dat.working)

```

#########################
*More Dataset Functions*
#########################

Further Summaries of Dataset
```{r}
#Located in install.packages() function
install.packages("psych")
library(psych)


##################
#Describe Function
##################
#describe() also provides us a clean summary/information on the variables in the dataset
describe(Excel.dat.working)
describe(Excel.dat.working, na.rm = TRUE)
  #describe(dataset, na.rm = TRUE/FALSE)


#Note: More information provided on describe function here: https://www.rdocumentation.org/packages/psych/versions/2.0.12/topics/describe
```

Describe By does the same function as describe but it can separate the datasets out by groups...

    - The PSYC200 dataset contains variables about GPA and variety of variables that affect GPA
    - Ex: # Hours students sleep, Note Mode (paper notes or computer notes), etc.
    
```{r}
#####################
#Describe By Function
#####################
#describe.by() provides us a clean summary/information on the variables in the dataset FOR A SPECIFIC GROUP

# Let's look at the means of test scores for each condition

describe.by(Excel.dat.working, Excel.dat.working$`Note Taking Modality (paper =1 or computer =2)`)
  #describe(dataset, dataset$group)
  

#Note3: More information provided here: https://www.rdocumentation.org/packages/psych/versions/2.0.12/topics/describeBy
```
