---
title: "R Module 3 Demonstration"
author: "David Xiao"
date: '2023-03-12'
output: html_document
---

```{r}
###################
#Before You Begin#
###################

- #Please download the updated version of the CSV STAT200 Dataset in the class Canvas Page, and then upload it to the BSOS R Cloud
- #We modified them so the variable names are shorter!
- #Follow the exact same step of download the file from the Class Canvas Page and Upload      them to the BSOS R Cloud like the last R Assignment 
      - "STAT200 Data Update.csv"
```


#########################################
#########################################
*Section One*
*Function Review & Referencing Variables*
#########################################
#########################################

- We will start this lesson by using the class dataset we used in Lesson 3
- Let's start by using the csv class data file


#A: Let's begin by opening/loading Dataset
```{r}
#First, you have to download the STAT 200 Data Update from the Class Canvas Page, and then Upload them to the BSOS R Cloud
library(readr)
library(readxl)
dat.org <- read.csv("STAT200 Data Update.csv")
dat.work <- dat.org

```

#B: Lets review related functions
```{r}
#Colnames() 
colnames(dat.work) #See all the variable names in dataset

#General Functions
str(dat.work) #See structure of all variables in dataset
summary(dat.work) #See numerical/character/factor summary of all variables in dataset

#Head/Tail
head(dat.work) #see first few rows
tail(dat.work) #see last few rows

#View
View(dat.work) #View entire dataset
#This is one of the only function that we will use the Capitalized "View"
```


#C: Now lets select specific Variables in the Dataset
```{r}
#This will help us visualize what we will do, it will show the headings, variables and a sample of the first few data. Especially helpful when you have 100,000 rows.  
head(dat.work)

#What if we just want to reference ONE of the variable. You would need the name of the data-set first, a $, and then the variable name.
#This is how to Extract a specific variable via $
dat.work$Sleep
dat.work$Motivation
dat.work$Scores

#This is how to Extract a specific variables via the bracket "[ ]" Indexing, refers to specific dimension of the data-sets. Data-frames tend to have two dimensions (the row and the column)

#COMMA will be used to separate out the dimensions
#Note: [Row, Column]
#Before the comma refers to "rows", and after the comma refers to "columns". 
#If nothing comes before the comma in the first space, you will be specifying all rows
dat.work[,2]
dat.work[,3]
#Colon: row 2 to 3
dat.work[,2:3]
dat.work[,3:4]
dat.work[,2:4]

```

#D: Now lets select Specific rows in the Dataset
```{r}
#Help Visualize what we will do... 
head(dat.work)

#Extracting specific rows (Row 1, Row 5 & Row 1:5)
dat.work[1,]
dat.work[5,]
dat.work[1:5,]
```

#E: Selecting Specific Variables Columns AND rows in the Dataset
```{r}
#Help Visualize what we will do... 
head(dat.work)

#Review: Here we can see that a [1,] indicates a row & indicates a column [,1]
#[row, columns]
dat.work[1,]
dat.work[,1]


#####################################
#Extracting specific variables & rows
#####################################

#Here we extract row 1 and column 1
dat.work[1,1] #we are referencing the first value of the row and first column
dat.work[1,2]
dat.work[2,4]
dat.work[5,5]
dat.work[34,6]

#Here we extract row 1 to 3 & column 1 to 3
# ":" means THROUGH
dat.work[1:3, 1:3]
dat.work[1:4, 1:4]

#Here we extract row 2 to 5, and 3 to 6 - be sure to stay in range.
dat.work[2:5, 3:6]
dat.work[1:50, 4:5]

```


############################################
############################################
*Section Two*
*Extracting & Building Vectors / Dataframes*
############################################
############################################

#A: Creating Vectors from Variables in Datasets
```{r}
#Help Visualize what we will do... 
head(dat.work)

#Now we access Student Sleep variable in the dataset, I am going to call this data-set SLEEP
Sleep <- dat.work$Sleep
Sleep

#Next we access Student Motivation in the dataset, I am going to call this data-set MOTIVATION
Motivation <- dat.work$Motivation
Motivation

#Now we access student Exam Scores in the dataset, I am going to call this data-set SCORES
Scores <- dat.work$Scores
Scores
```

#B: Selecting specific Parts of Vectors
```{r}
#We can also use the [row,column] function for specific values in a vector.

#Note we can only pull the rows since this is a vector and NOT a dataframe (so this only has one dimension), so we do not need the comma. Review ":" implies THROUGH

#First 3 Rows in Sleep
Sleep[1:3]
Sleep[2:7]

#First 3 Rows in Motivation
Motivation[1:3]

#First 3 Rows in Scores
Scores[1:3]


```

#C: Creating New Vectors Part 1


    - Here we will do some basic arithmetic using RStudio, except now we will reference specific variables in the dataset using the $
  
```{r}
############
#New Vectors
############
#Creating Vectors from the Dataset (and doing Math)

dat.work$Sleep
Sleep.DividedbyTwo <- dat.work$Sleep/2
Sleep.DividedbyTwo

Motivation.MultipliedbyTwo <- dat.work$Motivation*2
Motivation
Motivation.MultipliedbyTwo

Scores.Squared <- dat.work$Scores^2
Scores
Scores.Squared

Scores.Plus5<-dat.work$Scores+5
Scores
Scores.Plus5

Sleep
Motivation
Scores

Combining.Vectors <- dat.work$Sleep + dat.work$Motivation + dat.work$Scores
Combining.Vectors


####################
#Viewing New Vectors
####################
head(Sleep.DividedbyTwo)
head(Motivation.MultipliedbyTwo)
head(Scores.Squared)
head(Combining.Vectors)
```


#D: Creating New Vectors Part 2

    - We can create these same vectors by pulling them directly out of the dataset
    - Here we do this; and then select only the first 3 rows/rows of data
    
```{r}
############
#New Vectors
############
#Creating Vectors from the Dataset (first 3 rows/rows only)

First3.Sleep <- dat.work$Sleep[1:3]

First3.Motivation <- dat.work$Motivation[1:3]

First3.Scores <- dat.work$Scores[1:3]

#You can do the same with any of the data-sets


####################
#Viewing New Vectors
####################
First3.Sleep
First3.Motivation
First3.Scores

Combining.Vectors3<-First3.Sleep+First3.Motivation+First3.Scores
Combining.Vectors3
```



###############
*Section Three A*
*Subsetting*
###############

#Review Original Dataset Structure & Review Factors

    - In this section we will now "subset" specific variables and values in RStudio
```{r}
#Review Structure of Original Data
str(dat.org)

#Change Statistics Anxiety Variable from Numeric to Factor
dat.work$Stat.Anxiety <- as.factor(dat.work$Stat.Anxiety)
    #1 - Low Anxiety
    #2 - Medium Anxiety
    #3 - High Anxiety


#Checking if the change worked!
str(dat.org$Stat.Anxiety) #Original --> Numeric
str(dat.work$Stat.Anxiety) #Working --> Factor

levels(dat.work$Stat.Anxiety) #The level function will show us each of the factors of the variable
```

#Factor Subsetting
```{r}
#We briefly discussed Factors last week with the describeby() function
#DO NOT FORGET the PSYC Function, 
library(psych)
describeBy(dat.work, dat.work$Stat.Anxiety)

    #Here we see 3 datasets which reflect a summary of the variables per Statistics Anxiety Factor
          #1 - Low Anxiety #Mean = 80.33
          #2 - Medium Anxiety #Mean = 89.73
          #3 - High Anxiety #Mean = 76.43

```

#Specific method to Subset Variables by Factors

    - Here we look at exam scores by each factor of Statistics Anxiety by directly referencing the factor we are interested in using [ ] and then ==
    - You can see this below.
    
```{r}
##############################################
#Subset factors directly using the [] function
##############################################
#Just the exam Scores for Students with Low Stat Anxiety (Stat Anxiety == 1)
dat.work$Scores[dat.work$Stat.Anxiety == 1]

#Just the exam Scores for Students with Medium Stat Anxiety (Stat Anxiety == 2)
dat.work$Scores[dat.work$Stat.Anxiety == 2]

#Just the exam Scores for Students with High Stat Anxiety (Stat Anxiety == 3)
dat.work$Scores[dat.work$Stat.Anxiety == 3]


#########################################
#Mean of Scores per Each Subsetted Factor
#########################################
#Mean of Exam Scores for Students with Low Stat Anxiety (Stat Anxiety == 1)
mean(dat.work$Scores[dat.work$Stat.Anxiety == 1])

#Mean of Exam Scores for Students with Medium Stat Anxiety (Stat Anxiety == 2)
mean(dat.work$Scores[dat.work$Stat.Anxiety == 2])

#Mean of Exam Scores for Students with High Stat Anxiety (Stat Anxiety == 3)
mean(dat.work$Scores[dat.work$Stat.Anxiety == 3])

  #Note:Compare these means with the ones provided in the describe.by() function!
        #1 - Low Anxiety #Mean = 80.33
        #2 - Medium Anxiety #Mean = 89.73
        #3 - High Anxiety #Mean = 76.43

```

#We can use these subsets when we examine correlations for different groups! This is very useful when we do data analysis.
```{r}
#Correlation Test between all of the Sleep & Test Scores
cor1 <- cor.test(dat.work$Scores, dat.work$Sleep)
cor1

"Here we see the significant correlation because Exam Scores & the amount of sleep are correlated - as one may expect."
```

#We can use these subsets when we examine correlations for different groups! This is very useful when we do data analysis.
```{r}
#Correlation Test between Sleep & Test Scores for people who did Stat Anxiety Low
cor2<-cor.test(dat.work$Scores[dat.work$Stat.Anxiety ==1], dat.work$Sleep[dat.work$Stat.Anxiety==1])
cor2

#Correlation Test between Sleep & Test Scores for people who did Stat Anxiety High
cor3 <- cor.test(dat.work$Scores[dat.work$Stat.Anxiety == 3], dat.work$Sleep[dat.work$Stat.Anxiety == 3])
cor3


###################
#Note: Stat Anxiety
###################
  #1 = Low Anxiety
  #2 = Medium Anxiety
  #3 = High Anxiety

```

*Other Logical Operators for Subsetting Exist and are briefly reviewed below*

```{r}
#########
#Equal to
#########
#Equal to 1 (Low Anxiety)
dat.work$Scores[dat.work$Stat.Anxiety == 1]
mean(dat.work$Scores[dat.work$Stat.Anxiety == 1])


#############
#Not Equal to
#############
#Not Equal to 1 (Low Anxiety)
#The "!" infront of the equal sign means NOT equal
dat.work$Scores[dat.work$Stat.Anxiety !=1]

mean(dat.work$Scores[dat.work$Stat.Anxiety != 1])
#This code implies the mean for the scores for anxiety group for 2 or 3.
```

###############
*Section Three B*
*Subsetting*
###############


PART 2 #AND Logical Operator
```{r}
##################
#Preparing Dataset
##################
#We are going to examine the Note_Type_T1 Variable: whether a student took notes by hand or computer
#First Need Converting Note Taking Type at Time Period 1
dat.work$Note_Type_T1 <- as.factor(dat.work$Note_Type_T1)

#Viewing to See if it Worked
str(dat.work$Note_Type_T1)
levels(dat.work$Note_Type_T1)
    #Paper = 1
    #Computer = 2



#####################
#AND LOGICAL OPERATOR
#&
#####################

#Exam Scores for Statistics Anxiety Low (1) & Note Taking Mode Paper (1)
dat.work$Scores[dat.work$Stat.Anxiety == 1 & dat.work$Note_Type_T1 == 1]

#The Mean
mean(dat.work$Scores[dat.work$Stat.Anxiety == 1 & dat.work$Note_Type_T1 ==1])

#Length (How many rows qualified)
length(dat.work$Scores[dat.work$Stat.Anxiety == 1 & dat.work$Note_Type_T1 == 1])

#Exam Scores for Statistics Anxiety High (3) & Note Taking Mode Computer (2)
dat.work$Scores[dat.work$Stat.Anxiety == 3 & dat.work$Note_Type_T1 == 2]

#The Mean
mean(dat.work$Scores[dat.work$Stat.Anxiety == 3 & dat.work$Note_Type_T1 == 2])

#Length (How many rows qualified)
length(dat.work$Scores[dat.work$Stat.Anxiety == 3 & dat.work$Note_Type_T1 == 2])
```



##############################################
*There are also many other logical operators!*
##############################################

#Less than / Less than or equal to Logical Operator
```{r}
############
#Less than <
############
#Now we are going to look at the sleep variable again
dat.work$Scores[dat.work$Sleep < 70]

#The Mean    
mean(dat.work$Scores[dat.work$Sleep < 70])
    
#Length (How many rows qualified)
length(dat.work$Scores[dat.work$Sleep < 70])


#########################
#Less than or equal to <=
#########################
dat.work$Scores[dat.work$Sleep <=70]

#The Mean    
mean(dat.work$Scores[dat.work$Sleep <= 70])
    
#Length (How many rows qualified)
length(dat.work$Scores[dat.work$Sleep <= 70])
```


#Greater than / greater than or equal to Logical Operator
```{r}
###############
#Greater than >
###############
dat.work$Scores[dat.work$Sleep >70]

#The Mean
mean(dat.work$Scores[dat.work$Sleep > 70])
    
#Length (How many rows qualified)
length(dat.work$Scores[dat.work$Sleep > 70])


############################
#Greater than or equal to >=
############################
dat.work$Scores[dat.work$Sleep >= 70]

#The Mean
mean(dat.work$Scores[dat.work$Sleep >= 70])
    
#Length (How many rows qualified)
length(dat.work$Scores[dat.work$Sleep >= 70])
```


###################################
*More Information is Provided Here*
###################################

#https://www.statmethods.net/management/operators.html
#https://www.statmethods.net/management/subset.html


```


